--- 
title: "Juan Marin, Santiago Hurtado, Andres Parejo"
author: "Yihui Xie"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: rstudio/bookdown-demo
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Prerequisites

Placeholder



<!--chapter:end:index.Rmd-->

# Introduction {#intro}

You can label chapter and section titles using `{#label}` after them, e.g., we can reference Chapter \@ref(intro). If you do not manually label them, there will be automatic labels anyway, e.g., Chapter \@ref(methods).

Figures and tables with captions will be placed in `figure` and `table` environments, respectively.

```{r nice-fig, fig.cap='Here is a nice figure!', out.width='80%', fig.asp=.75, fig.align='center'}
par(mar = c(4, 4, .1, .1))
plot(pressure, type = 'b', pch = 19)
```

Reference a figure by its code chunk label with the `fig:` prefix, e.g., see Figure \@ref(fig:nice-fig). Similarly, you can reference tables generated from `knitr::kable()`, e.g., see Table \@ref(tab:nice-tab).

```{r nice-tab, tidy=FALSE}
knitr::kable(
  head(iris, 20), caption = 'Here is a nice table!',
  booktabs = TRUE
)
```
```{r}
data <- read.csv("bank.csv",sep=";",stringsAsFactors = TRUE)

head(data,n=5)

```
You can write citations, too. For example, we are using the **bookdown** package [@R-bookdown] in this sample book, which was built on top of R Markdown and **knitr** [@xie2015].

<!--chapter:end:Parejo.Rmd-->

```{r}
datos <- read.csv("bank.csv",sep=";")
head(datos, n=5)
dim(datos)
colnames(datos)
str(datos)

library(tidyverse)
 datos %>% 
  select(-day, -month, -previous) -> df
  head(df, n=5)
  df %>% 
  summarise(across(everything(),~ sum(is.na(.)), .names = "NA_{.col}"))
  tabla_y <- df %>% 
  count(y, name = "Frecuencia") %>% 
    mutate(Porcentaje = round(Frecuencia/sum(Frecuencia)*100,1),
    Etiqueta = paste0(Frecuencia, " (", Porcentaje, "%)"))
  tabla_y %>% 
  ggplot(aes(x = factor(y), y = Frecuencia)) +
  geom_col(fill = "purple", width = 0.6) +
  geom_text(aes(label = Etiqueta), vjust = -0.5, size = 3) +
  facet_grid(~ "Distribucion interes de deposito a plazo fijo (no/yes)") + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.15))) +
  labs(x = "Interes por el deposito a plazo fijo", y = "Numero de clientes (Porcentaje)") +
  #coord_flip() +
  theme_bw(base_size = 14) +
  theme(
    plot.title = element_blank(),
    strip.background = element_rect(fill = "gray80", color = NA),
    strip.text = element_text(face = "bold"),
    panel.grid.major.y = element_blank(),
    axis.text.y = element_text(size = 8)  
  )  
#falta interpretacion de la tabla de la variable objetivo
df %>% 
  summarise(
    n = length(age),
    media = mean(age),
    desviacion = sd(age),
    mediana = median(age),
    minimo = min(age),
    maximo = max(age),
    Q1 = quantile(age, 0.25),
    Q3 = quantile(age, 0.75),
    IQR = IQR(age)) %>%
  mutate(variable = "age") -> variable_age

df %>% 
  summarise(
    n = length(balance),
    media = mean(balance),
    desviacion = sd(balance),
    mediana = median(balance),
    minimo = min(balance),
    maximo = max(balance),
    Q1 = quantile(balance, 0.25),
    Q3 = quantile(balance, 0.75),
    IQR = IQR(balance)) %>% 
  mutate(variable = "balance") -> variable_balance

df %>% 
  summarise(
    n = length(duration),
    media = mean(duration),
    desviacion = sd(duration),
    mediana = median(duration),
    minimo = min(duration),
    maximo = max(duration),
    Q1 = quantile(duration, 0.25),
    Q3 = quantile(duration, 0.75),
    IQR = IQR(duration)) %>% 
  mutate(variable = "duration")-> variable_duration

df %>% 
  summarise(
    n = length(campaign),
    media = mean(campaign),
    desviacion = sd(campaign),
    mediana = median(campaign),
    minimo = min(campaign),
    maximo = max(campaign),
    Q1 = quantile(campaign, 0.25),
    Q3 = quantile(campaign, 0.75),
    IQR = IQR(campaign)) %>% 
  mutate(variable = "campaign")-> variable_campaign

df %>% 
  summarise(
    n = length(pdays),
    media = mean(pdays),
    desviacion = sd(pdays),
    mediana = median(pdays),
    minimo = min(pdays),
    maximo = max(pdays),
    Q1 = quantile(pdays, 0.25),
    Q3 = quantile(pdays, 0.75),
    IQR = IQR(pdays)) %>% 
  mutate(variable = "pdays")-> variable_pdays

bind_rows(variable_age, variable_balance, variable_duration,variable_campaign,variable_pdays) %>%
  select(variable, everything())
```

<!--chapter:end:Marin.Rmd-->


# Análisis exploratorio bivariado.

Placeholder



<!--chapter:end:Hurtado.Rmd-->

